---
title: Mackrics
output:
  litedown::html_format:
    meta:
      css: ["index-style.css", "styles.css", "https://prismjs.catppuccin.com/mocha.css"]
---

::: {.callout-warning}

**This is the development version of my website**

Please go to <https://mackrics.com/> for the main version.
:::

```{r, setup}
#| echo = FALSE,
#| message = FALSE

source(here::here("packages.R"))
source("./R/gen_rss_feed.R")
```

# Welcome to my personal blog!

I’m an economist (the original data science) exploring open source software and hardware.
Here are some programs I've written:

- [Ricethemes](https://github.com/Mackrics/ricethemes): an R package with cool
  ggplot2 themes.
- [Neoslime](https://github.com/Mackrics/neoslime): a neovim plugin to send
  code from the editor to a neovim terminal.


# Latest blog posts

```{r}
#| results = 'asis',
#| echo = FALSE

gen_blogroll <- function(nrow) {
  fs::dir_ls("blog", regex = "*.Rmd") |>
  data.table(path = _) |>
  data.table::copy() |>
  _[path != "blog/index.Rmd"] |>
  _[, let(
      params = lapply(path, rmarkdown::yaml_front_matter)
    )] |>
  _[, let(
        date = as.Date(purrr::map_chr(params, \(x) x[[ "date" ]])),
        title = purrr::map_chr(params, \(x) x[[ "title" ]]),
	link = stringr::str_replace(path, "Rmd", "html")
  )] |>
  _[,let(
    print = purrr::pmap_chr(
      list(date, title, link), 
      \(date, title, link) glue::glue("\n\n**{date}**: [{title}]({link})\n\n")))
  ][order(-date), print] |>
  head(5) |>
  cat()
}
gen_blogroll(5)
```

[View all](./blog/index.html)
