---
title: Introducing `{ricethemes}`
date: 2023-09-21
tags: "R"
---

I've written a R package with popular desktop themes for `{ggplot2}`. The idea
is that the package will enable analysts to explore data using their preferred
desktop theme and have a coherent development enviorment. For now, I've added
the [Catppuccin](https://github.com/catppuccin/catppuccin) theme with four
flavors: `latte`, `frappe`, `macchiato`, and `mocha`. I plan on adding more
themes in the future, depending on the demand.

I recommend using the [`pak`](https://pak.r-lib.org/) package to install
ricethemes.

```r
pak::pkg_install("mackrics/ricethemes")
```

The package contain functions to obtain hex codes for the various colors,
themes, and functions to display the colors.


```{r}
#| echo = FALSE,
#| message = FALSE

source(here::here("packages.R"))

list(ctp_mocha, ctp_macchiato, ctp_frappe, ctp_latte) |>
  lapply(show_ctp_theme) |>
  patchwork::wrap_plots() +
  plot_annotation() & 
  theme(plot.margin = margin(0, 0, 0, 0))  # Set overall plot margins to zero
```

The classic `cars` correlation plot looks something like this for the four
flavors:

```{r}
#| echo = FALSE

list(ctp_mocha, ctp_macchiato, ctp_frappe, ctp_latte) |>
  lapply(\(ctp_theme)
    ggplot(cars) +
    aes(speed, dist) +
    geom_point(color = {{ ctp_theme }}[["yellow"]]) +
    geom_smooth(
      method = "lm",
      se = FALSE,
      formula = "y ~ x",
      color = {{ ctp_theme }}[["peach"]]
    ) +
    theme_ctp({{ ctp_theme }}) +
    labs(x = "Speed", y = "Distance")
  ) |>
  patchwork::wrap_plots() +
  plot_annotation() & 
  theme(plot.margin = margin(0, 0, 0, 0))  # Set overall plot margins to zero
```

Any feedback or request is much appreciated, preferably as a GitHub issue in
[this repository](https://github.com/Mackrics/ricethemes/issues).
