---
title: Introducing `{ricethemes}`
date: 2023-09-21
date-format: iso
---

I've written a R package with popular desktop themes for `{ggplot2}`. The idea
is that the package will enable analysts to explore data using their preferred
desktop theme and have a coherent development enviorment. For now, I've added
the [Catppuccin](https://github.com/catppuccin/catppuccin) theme with four
flavors: `latte`, `frappe`, `macchiato`, and `mocha`. I plan on adding more
themes in the future, depending on the demand.

I recommend using the [`pak`](https://pak.r-lib.org/) package to install
ricethemes.

```r
pak::pkg_install("mackrics/ricethemes")
```

For now, the package contains functions to obtain hex codes for the various colors,
themes, and functions to display the colors.

```{r}
#| panel: tabset
#| output: asis
#| fig-align: center
#| out-width: "100%"
library(ricethemes)
library(ggplot2)
library(tibble)

plotlist <-
  tibble(ctp_mocha, ctp_macchiato, ctp_frappe, ctp_latte) |>
  lapply(show_ctp_theme)

for (i in seq(1, length(plotlist))) {
  cat("\n\n#", names(plotlist)[[ i ]], "\n")
  print(plotlist[[ i ]])
}
```

The classic `cars` correlation plot looks something like this for the four
flavors:

```{r}
#| panel: tabset
#| output: asis
#| fig-align: center
#| out-width: "100%"
plotlist2 <-
  tibble(ctp_mocha, ctp_macchiato, ctp_frappe, ctp_latte) |>
  lapply(\(ctp_theme)
    ggplot(cars) +
    aes(speed, dist) +
    geom_point(color = {{ ctp_theme }}[["yellow"]]) +
    geom_smooth(
      method = "lm",
      se = FALSE,
      formula = "y ~ x",
      color = {{ ctp_theme }}[["peach"]]
    ) +
    theme_ctp({{ ctp_theme }}) +
    labs(x = "Speed", y = "Distance")
  )

for (i in seq(1, length(plotlist2))) {
  cat("\n\n#", names(plotlist2)[[ i ]], "\n")
  print(plotlist2[[ i ]])
}
```

Any feedback or request is much appreciated, preferably as a GitHub issue in
[this repository](https://github.com/Mackrics/ricethemes/issues).
